Integer-N PFD/CP-based PLL - System-Level simulation

### 2025-05-08 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Removing Laboratories and Examples folder and unifying everything in the folder CaseStudies.

---
title: Integer-N PFD/CP-based PLL - System-Level simulation
subtitle: Laboratory 01 - Week 01-09
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Este trabalho apresenta duas simulações para um Integer-N PFD/CP-based PLL a nível de sistema. Uma simulação é no domínio do tempo e a outra no domínio da fase. Foram estabelecidas restrições para a operação e são apresentados os passos até a concepção dos sistemas para a simulação. Por fim, os resultados são analisados quanto ao tempo de resposta e comparados ao que foi previsto com base na teoria.
keywords: 
  - Phase-locked loop
  - Sistema de controle
  - PLL Integer-N
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introdução

@@

This report shows the development of a Integer-N PFD/CP-based PLL with a second-order loop filter. The loop filter is designed so that the loop bandwidth has a damping constant $\zeta=$ `{python} zeta`, and a 3-dB bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz.

The intention here is to answer the two following questions:

- What is the maximum frequency step at the input such that the theory so far is still able to predict the behavior of the system?

- Provided this condition is met, how long will it take the system to settle from a frequency step?

To calculate the natural frequency of the loop it's used
$$
\begin{split}
\omega_{3\text{dB}}&=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}\implies\\
\omega_n&=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
\end{split}
$$

@@

So that the $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s. With this value, it's possible to calculate the instant that occurs the maximum value of the phase error, using the following equation

@@

Then, applied in the following equation, it's gotten the maximum normalized phase error to a frequency step, $\theta_{e_\text{m,n}}$, from

@@

Resulting in the maximum normalized phase error to a frequency step is about `{python} round(theta_e_m_n, 2)`. This value can be applied to the non-normalized equation of the phase error
$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

As the value of interest is that step frequency, $\Delta\omega_m$, that extrapolates the maximum value of phase error, $\theta_{e_m}$, that normally is $\theta_{e_m}=2\pi$. To calculate this, it's used $\theta_{e_\text{m,n}}$, as follows

@@

Applying the already obtained values, it results in $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s, or `{python} format(step_m_lin)`Hz.

So the maximum frequency step at the input such that the theory so far is still able to predict the behavior of the system is `{python} format(step_m_lin)`Hz. If the frequency step is any larger than this, then the PLL will lose lock and cycle slip, which is the

At other side, the settling time, $t_s$, can be calculated as

@@

resulting in a settling time of $t_s=$`{python} format(t_s)`s.

@@

>>>>

---
title: Exploring the Validity Range of PLL Analytical Models
subtitle: Case Study - Example 3.2 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the analytical development and performance estimation of a second-order Integer-N Phase-Locked Loop (PLL) based on a Phase-Frequency Detector (PFD) and a Charge Pump (CP). The loop filter is designed to meet specified damping and bandwidth characteristics. The analysis determines the maximum allowable frequency step at the input that ensures the PLL remains in lock, based on the system's maximum normalized phase error. Additionally, the settling time following such a frequency step is estimated using classical control theory. The results provide useful insights into the dynamic behavior of the loop, serving as a theoretical benchmark for future simulation and implementation efforts.
keywords: 
    - Integer-N PLL
    - Loop Dynamics
    - Frequency Step
    - Settling Time 
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introduction

@@

This report presents the development of an Integer-N Phase-Frequency Detector/Charge Pump-based Phase-Locked Loop (PFD/CP-PLL) with a second-order loop filter. The loop filter is designed so that the loop bandwidth has a damping factor $\zeta=$ `{python} zeta`, and a 3-dB bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. The example analyzed in this report is based on the case presented in @rogers2006IntegratedCircuitDesign.

## Objective

The objective of this study is to answer the following two questions:

- What is the maximum frequency step at the input such that the current theoretical model remains valid in predicting the system's behavior?
- Assuming this condition is satisfied, how long will the system take to settle following a frequency step?

# Analysis

## Loop Natural Frequency Calculation

To compute the natural frequency of the loop, the following equation is used:

$$
\begin{split}
\omega_{3\text{dB}}&=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}\implies\\
\omega_n&=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
\end{split}
$$

@@

Thus, the natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s.

## Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

@@

## Maximum Normalized Phase Error

Using the value of $t_m$, the maximum normalized phase error for a frequency step, $\theta_{e_\text{m,n}}$, is given by:

@@

As a result, the maximum normalized phase error is approximately `{python} round(theta_e_m_n, 2)`.

## Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the maximum phase error does not exceed $\theta_{e_m}=2\pi$. Using the normalized result:

@@

Substituting the calculated values yields:

- $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s
- $\Delta f_m=$ `{python} format(step_m_lin)`Hz

Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)` Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping.

## Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

@@

This results in a settling time of $t_s=$ `{python} format(t_s)`s.

@@

# Conclusion

This report provides a theoretical analysis of a second-order PFD/CP-based Integer-N PLL. The maximum tolerable frequency step for which the linear model remains valid was derived, and the corresponding settling time was estimated. These results serve as a foundation for further simulation and validation studies.

### 2025-05-08 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd
Removing Laboratories and Examples folder and unifying everything in the folder CaseStudies.

---
title: Integer-N PFD/CP-based PLL - System-Level simulation
subtitle: Laboratory 01 - Week 01-09
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Este trabalho apresenta duas simulações para um Integer-N PFD/CP-based PLL a nível de sistema. Uma simulação é no domínio do tempo e a outra no domínio da fase. Foram estabelecidas restrições para a operação e são apresentados os passos até a concepção dos sistemas para a simulação. Por fim, os resultados são analisados quanto ao tempo de resposta e comparados ao que foi previsto com base na teoria.
keywords: 
  - Phase-locked loop
  - Sistema de controle
  - PLL Integer-N
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introdução
The intention in this report is the designing of an integer-N synthesizer that operates at `{python} format(f0)`Hz to `{python} format(f)`Hz. The synthesizer must to settle for a frequency step in a settling time of, $t_s=$ `{python} format(t_s)`s. The channel spacing is `{python} format(f_ref)`Hz and as it's using an integer-N synthesizer, the channel space is equal to the frequency reference.

Using a equation discussed in the previous report, it's possible to get the natural frequency, $\omega_n$,

@@

Using an arbitrary value for $\zeta=$ `{python} zeta`. With that constant, the natural frequency would be $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, or $f_n=$ `{python} format(f_n)`Hz.

Starting from $omega_n$ and $zeta$, it's possible to get the 3-dB bandwidth, $\omega_{3\text{dB}}$ as follows,

@@

Using that relation, it results in $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz.

How the relation between $f_\text{ref}\gg f_{3\text{dB}}$, thus the system isn't affected for the discretization process (sampling). So the continuous analysis is useful for this example.

@@

As it's an integer-N and the reference frequency $f_\text{ref}=$ `{python} format(f_ref)`Hz, the so $N$ should be from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## VCO

@@

To the VCO, it's needed to cover a tuning range of  `{python} format(f_range)`Hz.

Due to process variations, it is necessary to ensure that the VCO has an additional tuning margin. Therefore, a total tuning range of approximately `{python} round(tuning_range_percent, 0)`%, of the center frequency, around `{python} format(tuning_range)`Hz, is recommended. And assuming a voltage supply of `{python} format(V)`V, a VCO constant of `{python} format(Kvco)`Hz/V `{python} format(Kvco_ang)`rad/s/V is expected. So the free running frequency of the VCO should be in minimum frequency. The middle of the frequencies, $f_c$,

@@

So it results in $f_c=$ `{python} format(f_c)`Hz. The minimum frequency is $f_\text{min}=$ `{python} format(f_min)`Hz and the maximum is `{python} format(f_max)`Hz. The free running frequency will be equals to $f_\text{VCO}=f_\text{min}$. So, if it's demanded a frequency of `{python} format(f_example_1, 3)`Hz, for example, it uses the following equation to determine the voltage of control, $V_c$

@@

So the voltage control will be $V_c=$ `{python} format(V_c_example_1)`V.

@@

As it was commented before, the continuous-time analysis can be useful in this system. So the already known equations for calculation of the second-order loop filter. For the first capacitor, $C_1$, its value can be obtained as

@@

The charge pump current, $I_\text{CP}$, is another unknown variable in this equation, so it's arbitraryly supposed a value of $C_1=$ `{python} format(C_1)`F, so the charge pump current should be $I_\text{CP}=$ `{python} format(I_CP_ang)`A. Then the resistor can be determined as

@@

Thus the resistor will be $R=$ `{python} format(R_1)`$\Omega$.

Using the common realtion, it's possible to find the second capacitor, $C_2$, for the second-order loop filter,

@@

Resulting in $C_2=$ `{python} format(C_2)`F.

>>>>

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of robust and efficient frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introduction

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

# Design Methodology

## Frequency Settling and Loop Dynamics

To meet the settling time requirement, the natural frequency $\omega_n$ of the loop must be determined. Using an equation discussed in a previous report, the relationship between settling time, damping factor $\zeta$, and natural frequency is given by:

@@

Assuming an arbitrary value of $\zeta=$ `{python} zeta`, the resulting natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, which corresponds to a frequency $f_n=$ `{python} format(f_n)`Hz.

From the natural frequency and damping factor, the 3-dB bandwidth $\omega_{3\text{dB}}$ can be calculated as:

@@

This results in a bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. Given that $f_\text{ref}\gg f_{3\text{dB}}$, the discretization effects due to sampling are negligible, and a continuous-time analysis remains valid for this system.

@@

Since the architecture is integer-N and the reference frequency is $f_\text{ref}=$ `{python} format(f_ref)`Hz, the division ratio $N$ must range from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## Voltage-Controlled Oscillator (VCO)

@@

The VCO must span a tuning range of `{python} format(f_range)`Hz. To account for process variations, an additional tuning margin is considered. A total tuning range of approximately `{python} round(tuning_range_percent, 0)`% of the center frequency, or about `{python} format(tuning_range)`Hz, is recommended.

Assuming a supply voltage of `{python} format(V)`V, a VCO gain of `{python} format(Kvco)`Hz/V (or `{python} format(Kvco_ang)`rad/s/V) is expected. The free-running frequency of the VCO is set at the minimum frequency. The center frequency $f_c$ is given by:

@@

This results in $f_c=$ `{python} format(f_c)`Hz. Therefore, the minimum and maximum VCO frequencies are $f_\text{min}=$ `{python} format(f_min)`Hz and $f_\text{max}=$ `{python} format(f_max)`Hz, respectively. The free-running frequency is set as $f_\text{VCO}=f_\text{min}$.

For instance, to synthesize a frequency of `{python} format(f_example_1, 3)`Hz, the required control voltage $V_c$ is computed using:

@@

This yields a control voltage of $V_c=$ `{python} format(V_c_example_1)`V.

@@

As previously noted, continuous-time analysis is applicable. Standard second-order loop filter equations are used. The first capacitor $C_1$ can be calculated from:

@@

Since the charge pump current $I_\text{CP}$ is unknown, a value of $C_1=$ `{python} format(C_1)`F is assumed, leading to a calculated $I_\text{CP}=$ `{python} format(I_CP_ang)`A. The resistor $R_1$ is then given by:

@@

Thus, the resistor value is $R=$ `{python} format(R_1)`$\Omega$.

Using a common design heuristic, the second capacitor $C_2$ is:

@@

Which results in $C_2=$ `{python} format(C_2)`F.

# Conclusion

This report has outlined the key steps in designing an integer-N frequency synthesizer operating within the specified frequency range. A continuous-time model was adopted due to the large separation between the reference frequency and the loop bandwidth. Component values for the VCO and loop filter were estimated based on reasonable assumptions, providing a solid foundation for subsequent implementation and simulation.

### 2025-05-08 CaseStudies/Lab01_IntegerNSystemLevel/Report/article/article.qmd
Removing Laboratories and Examples folder and unifying everything in the folder CaseStudies.

@@

Além disso, a [Fig. @fig-phase-domain-system] apresenta dois outros ganhos que não foram muito apresentados, o ganho $A_0$ e $K_\theta$ ou $K_\text{phase}$. O ganho $A_0$ representa o ganho em corrente contínua e é assumido unitário nas simulações deste trabalho. O outro elemento de ganho, $K_\theta$, por sua vez é utilizado para quantizar a quantidade de corrente fornecida pelo charge pump durante um ciclo de operação. Se o charge pump conduziu pelo ciclo completo, a corrente média, $i_d=I$, se o charge pump conduziu por meio ciclo, $i_d=I/2$ e assim por diante. O que define o ângulo de condução do charge pump é a diferença de fase, $\theta_R-\theta_o$. Se a diferença for de $90^\circ$ ($\pi/2$), o carge pump vai conduzir por 1/4 do período e, portanto, $i_d=I/4$. Por isso, se tomarmos $\tau$ como o tempo de condução e $T$ o período, a definição da corrente média, $i_d$, é dada por @rogers2006IntegratedCircuitDesign

@@

@@

>>>>

Além dos ganho do VCO apresentado, este simples sistema apresentado aqui possui dois outros ganhos que não foram comentados até então; o ganho $A_0$ e $K_\theta$ ou $K_\text{phase}$. O ganho $A_0$ representa o ganho em corrente contínua e é assumido unitário nas simulações deste trabalho. O outro elemento de ganho, $K_\theta$, por sua vez é utilizado para quantizar a corrente média fornecida pelo charge pump durante um ciclo de operação. Se o charge pump conduziu pelo ciclo completo, a corrente média será $i_d=I$, se o charge pump conduziu por meio ciclo, $i_d=I/2$ e assim por diante. O que define o ângulo de condução do charge pump é a diferença (ou erro) de fase, $\theta_R-\theta_o$. Se a diferença for de $90^\circ$ ($\pi/2$), por exemplo, o carge pump vai conduzir por 1/4 do período e, portanto, $i_d=I/4$. Por isso, se tomarmos $\tau$ como o tempo de condução e $T$ o período, a definição da corrente média, $i_d$, é dada por @rogers2006IntegratedCircuitDesign

@@

@@

<!--

@@

-->
